---
title: "Lab 4"
#Max Ward
output: html_document
---

```{r,echo=FALSE,message = FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
babies <- read_csv("https://raw.githubusercontent.com/ervance1/Sp2018-Data-Science-repo/master/babies2a.dat")
babies <- rename(babies, bwtoz = `bwt/oz`)
babies <- na.omit(babies)
```
#Max
```{r,echo=FALSE,message = FALSE, warning=FALSE}
babies$mrace[babies$mrace == 1] <- 0
babies$mrace[babies$mrace == 2] <- 0
babies$mrace[babies$mrace == 3] <- 0
babies$mrace[babies$mrace == 4] <- 0
babies$mrace[babies$mrace == 5] <- 0
view(babies)
```
Quetion: Does the mothers race and or gestation effect the birth weight of the baby?
```{r,echo=FALSE,message = FALSE, warning=FALSE}
ggplot(data = babies, aes(x = gestation, y = bwtoz , color= as.factor(mrace))) +
  geom_point()+
  facet_wrap(~mrace)+
  geom_smooth(method = "lm", se= FALSE)+
  scale_color_manual(labels = c("White", "Mexican", "Black", "Asian", "Mix", "Unknown"), values = c("white", "blue", "yellow", "pink", "green", "red"))
```

```{r,echo=FALSE,message = FALSE, warning=FALSE}
fit_1  <- lm(bwtoz ~ gestation, data = babies)
summary(fit_1)
```
As we can see from the data the p-value is small so we can deterime the data is significant and it did not just happen by chance. 
```{r}
plot(bwtoz ~ gestation, data = babies, col = "blue")
abline(fit_1)
```

Answer: Based on the smaller graphs it does look like the race of the mom has an effect on the babies birth weight. As you can see some of the lines are much steeper than the others. For the gestation for all the races it looks like as the gestation is longer the baby is higher.

What I ddi:I graphed my data in a much easier way to read and I used the p-value to test if the data was significant. I also added a third variable to comapare all the data. 
